---
title: "FW 536 — Post-course exam"
author: "Your Name Here"
date: "`r Sys.Date()`"
output:
  html_document:
    self_contained: true
    toc: true
    toc_depth: 3
    toc_float: true
    code_folding: show
    theme: cosmo
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
```

# Instructions

This template matches the [Post-course exam](post_course_exam.html). Fill
in each section with your answer. Show R code where helpful; embed photos
of hand-written derivations where useful.

You have 90 minutes. You may use your notes, the course materials, and R.
No collaboration.

When done: **Knit** to HTML and upload the `.html` to Canvas.

---

# Part A — Probability fundamentals (6 points)

## A1. Bayes' rule applied (2 pts)

*Sensitivity 0.90, specificity 0.96, prior 0.10. Posterior probability given one positive?*

```{r}
sens <- 0.90; spec <- 0.96; fpr <- 1 - spec
prior <- 0.10
# (sens * prior) / (sens * prior + fpr * (1 - prior))
```

**Answer:**

## A2. Updating from two samples (2 pts)

```{r}
# update twice
```

**Answer:**

## A3. Joint probability with imperfect detection (2 pts)

*P(present) = 0.40; per-night p = 0.25; three nights. P(zero detections)?*

```{r}
# tip: condition on presence vs absence
```

**Answer:**

---

# Part B — Distributions (6 points)

## B1. Pick the distribution and why (2 pts)

*Counts with mean ≈ 5, variance ≈ 22.*

**Answer:**

## B2. Beta-Binomial (2 pts)

*Why does heterogeneity inflate the probability of "all-zero" deployments?*

**Answer:**

## B3. Continuous distribution with a bound (2 pts)

*Percent cover, strictly between 0 and 1, not from a count.*

**Answer:**

---

# Part C — Regression and GLMs (10 points)

## C1. Log-link biology (3 pts)

*Poisson regression slope on temperature = -0.18 per °C.*

**a. Multiplicative effect of 1 °C warming:**

```{r}
exp(-0.18)
```

**b. Temperature increase that halves expected counts:**

```{r}
-log(2) / -0.18
```

## C2. Logit-link biology (3 pts)

*Intercept = -3, slope = +0.05 per percent canopy.*

**a. Probability at canopy = 0:**

```{r}
plogis(-3)
```

**b. Canopy at which p = 0.5:**

```{r}
# solve -3 + 0.05 * x = 0
```

**c. Marginal effect at canopy = 60:**

```{r}
# beta * p * (1 - p) where p is predicted at canopy = 60
```

## C3. Overdispersion diagnosis (2 pts)

*Residual deviance 412 on 80 df.*

**Answer:**

## C4. Interaction interpretation (2 pts)

*Main temp = -0.10; gear[Trawl] = +0.30; temp:gear[Trawl] = +0.15. Temp effect on trawl?*

```{r}
# slope for trawl = -0.10 + 0.15
# exp of that
```

**Answer:**

---

# Part D — Mixed models / hierarchy (8 points)

## D1. Notation (2 pts)

Write the mixed-effects model with random intercept by site in math notation.

**Answer (use $math$ or display $$math$$):**

## D2. REML vs ML (3 pts)

**Answer:**

## D3. Random slopes (3 pts)

*Var(intercept) = 4.2, Var(slope) = 0.05, correlation = -0.6. Biological meaning?*

**Answer:**

---

# Part E — Maximum likelihood (6 points)

## E1. MLE by hand (3 pts)

*Bernoulli, n = 40, k = 12. Derive MLE of p (show the steps).*

**Derivation:**

(You may embed a photo of hand-written algebra here.)

```{r}
# numerical check: MLE should be k/n
12 / 40
```

## E2. LRT and profile-likelihood CI (3 pts)

**LRT (one sentence):**

**Profile-likelihood 95% CI (the "1.92" cutoff):**

---

# Part F — Bayes (8 points)

## F1. Conjugacy (3 pts)

*18 successes in 30 trials; prior Beta(2, 8). Posterior and mean?*

```{r}
# posterior: Beta(2 + 18, 8 + 12) = Beta(20, 20)
# mean: 20/40
```

**Answer:**

## F2. BCI vs HPDI (2 pts)

**Answer:**

## F3. Sketch a Nimble model (3 pts)

```r
# nimbleCode({
#   for (i in 1:n) {
#     log(lambda[i]) <- beta0 + beta1 * x[i]
#     y[i] ~ dpois(lambda[i])
#   }
#   beta0 ~ dnorm(0, sd = 10)
#   beta1 ~ dnorm(0, sd = 10)
# })
```

**Your answer (clean up / adjust the above as you like):**

---

# Part G — Applied case study (6 points)

## G1. From question to model

*20 streams × 5 surveys × 3 years of redd counts; survey-level covariates
(temperature, discharge); stream-level covariate (management designation,
3 levels). Question: does management affect redd abundance after
accounting for within-stream variation?*

**a. Distribution and link function:**

**b. Grouping structure (streams, years):**

**c. Adequate-fit check:**

**d. Frequentist or Bayesian, and why:**

---

# Submission checklist

- [ ] Filled in every section.
- [ ] Code chunks knit without error.
- [ ] Photos (if any) embedded via `self_contained: true`.
- [ ] Uploaded the `.html` to Canvas.
